---
title: "Mini-Project 02 - Advancing Housing Accessibility Across America"
author: "Student Analysis"
editor:
    mode: source
format:
    html:
        theme: cosmo
        toc: true
        toc-location: left
        toc-depth: 3
        toc-title: "Document Navigation"
        code-fold: true
        code-summary: "Display code"
        code-tools: true
        reference-location: margin
        citation-location: margin
        smooth-scroll: true
        page-layout: full
        grid:
          sidebar-width: 300px
          body-width: 900px
          margin-width: 250px
execute:
  message: false
  warning: false
---

## Introduction

::: {.callout-note icon=false}
## Project Overview

From the [STA 9750 Course Assignment Page](https://michael-weylandt.com/STA9750/miniprojects/mini02.html):

> "Housing affordability is one of the greatest challenges facing New York City today. ... In this mini-project, we will identify America's most "YIMBY" cities using a variety of census data sources and real estate indices. ... You will use the results of your analysis to lobby politicians in support of a federal YIMBY-incentive program. In particular, you will prepare a short policy brief-a sort of written elevator pitch-that is designed to help you find congressional representatives whose districts would benefit from adopting YIMBY-type policies and whose reelection prospects would be improved by sponsoring this type of bill."

:::

The objective of this analysis is to develop data analysis capabilities, analytical reasoning, and professional communication competencies. By leveraging R programming skills and analytical methods acquired throughout this course, this document will produce a mock policy brief demonstrating the significance of the YIMBY ("yes in my backyard") initiative using empirical data. 

This document builds upon methodologies employed in [Mini-Project 1](https://rcutsumpas-cloud.github.io/STA9750-2025-FALL/mp01.html), incorporating enhanced data processing techniques, developing visual representations to communicate data narratives, and consolidating intricate information into accessible outputs for diverse audience expertise levels.
---

### NIMBY vs YIMBY Policy Approaches

This analysis references specialized policy and industry terminology. The definitions provided below offer context for various terms and acronyms utilized throughout this document.

::: {.callout-warning icon=false}
## Essential Terminology

*NIMBY: *"Not In My Backyard" - resistance to proposed development and housing initiatives in residential areas that residents consider undesirable

*YIMBY: *"Yes In My Backyard" - advocacy movement supporting residential development projects within local communities 

*Housing Crisis*: ongoing deficit in affordable housing construction relative to population demand

*CBSA: *Core Based Statistical Area - metropolitan and micropolitan regions designated for statistical purposes

*NAICS Codes: *North American Industry Classification System - standardized classification framework for business categorization by sector and type

*Individual Income*: aggregate earnings for a single person from employment, Social Security, and supplementary benefits

*Industry Wages*: mean compensation provided to workers within a particular sector

*Rent Burden: proportion of renters allocating *more than 30% of monthly earnings toward housing and utility costs

*Severe Rent Burden: proportion of renters allocating *more than 50% of monthly earnings toward housing and utility costs

*Housing Growth*: expansion in housing availability, market demand, and property valuations across time

:::

<p style="text-align: left; font-size: 0.7em; color: #666; font-style: italic; margin-top: -10px;">
Sources: Britannica; Cambridge Dictionary; US Census Bureau; US Bureau of Economic Analysis; Investopedia; Neighborhood Data for Social Change
</p>

---

## Data Loading

This analysis employs publicly available datasets from the US Census Bureau's American Community Survey (ACS) and the Bureau of Labor Statistics. These repositories contain demographic, housing, and employment information from authoritative governmental agencies. Although these datasets may contain certain limitations including potential population undercounts or underrepresented demographic groups, the Census and BLS maintain comprehensive, high-quality nationwide data. These repositories additionally offer thoroughly documented and pertinent information that enhances credibility for this project's objective of securing policymaker endorsement.

```{r}
#| label: Census Dataset Loading
#| code-fold: true
#| code-summary: "Display US Census Bureau Loading Code"

if(!dir.exists(file.path("data", "mp02"))){
    dir.create(file.path("data", "mp02"), showWarnings=FALSE, recursive=TRUE)
}

verify_package <- function(pkg){
    pkg <- as.character(substitute(pkg))
    options(repos = c(CRAN = "https://cloud.r-project.org"))
    if(!require(pkg, character.only=TRUE, quietly=TRUE)) install.packages(pkg)
    stopifnot(require(pkg, character.only=TRUE, quietly=TRUE))
}

verify_package(tidyverse)
verify_package(glue)
verify_package(readxl)
verify_package(tidycensus)

retrieve_acs_data <- function(variable, geography="cbsa",
                              initial_year=2009, final_year=2023){
    filename <- glue("{variable}_{geography}_{initial_year}_{final_year}.csv")
    filename <- file.path("data", "mp02", filename)
    
    if(!file.exists(filename)){
        YEAR_RANGE <- seq(initial_year, final_year)
        YEAR_RANGE <- YEAR_RANGE[YEAR_RANGE != 2020] # Exclude 2020 - Survey not conducted (covid)
        
        COMPLETE_DATA <- map(YEAR_RANGE, function(yr){
            tidycensus::get_acs(geography, variable, year=yr, survey="acs1") |>
                mutate(year=yr) |>
                select(-moe, -variable) |>
                rename(!!variable := estimate)
        }) |> bind_rows()
        
        write_csv(COMPLETE_DATA, filename)
    }
    
    read_csv(filename, show_col_types=FALSE)
}

# Household income (annual)
HOUSEHOLD_INCOME <- retrieve_acs_data("B19013_001") |>
    rename(household_income = B19013_001)

# Monthly rent
MONTHLY_RENT <- retrieve_acs_data("B25064_001") |>
    rename(monthly_rent = B25064_001)

# Total population
TOTAL_POPULATION <- retrieve_acs_data("B01003_001") |>
    rename(population = B01003_001)

# Total number of households
TOTAL_HOUSEHOLDS <- retrieve_acs_data("B11001_001") |>
    rename(households = B11001_001)
```

```{r}
#| label: Housing Development Data Loading
#| code-fold: true
#| code-summary: "Display Census Housing Development Loading Code"

retrieve_development_permits <- function(initial_year = 2009, final_year = 2023){
    filename <- glue("housing_developments_{initial_year}_{final_year}.csv")
    filename <- file.path("data", "mp02", filename)
    
    if(!file.exists(filename)){
        PAST_YEARS <- seq(initial_year, 2018)
        
        PAST_DATA <- map(PAST_YEARS, function(yr){
            past_url <- glue("https://www.census.gov/construction/bps/txt/tb3u{yr}.txt")
                
            TEXT_LINES <- readLines(past_url)[-c(1:11)]

            REGION_LINES <- str_detect(TEXT_LINES, "^[[:digit:]]")
            REGION_CODE <- as.integer(str_sub(TEXT_LINES[REGION_LINES], 5, 10))

            AUTHORIZATION_LINES <- str_detect(str_sub(TEXT_LINES, 48, 53), "[[:digit:]]")
            AUTHORIZATIONS <- as.integer(str_sub(TEXT_LINES[AUTHORIZATION_LINES], 48, 53))
            
            data_frame(CBSA = REGION_CODE,
                       new_housing_units_permitted = AUTHORIZATIONS, 
                       year = yr)
        }) |> bind_rows()
        
        RECENT_YEARS <- seq(2019, final_year)
        
        RECENT_DATA <- map(RECENT_YEARS, function(yr){
            recent_url <- glue("https://www.census.gov/construction/bps/xls/msaannual_{yr}99.xls")
            
            temp_file <- tempfile()
            
            download.file(recent_url, destfile = temp_file, mode="wb")
            
            backup_function <- function(.func1, .func2){
                function(...){
                    tryCatch(.func1(...), 
                             error=function(e) .func2(...))
                }
            }
            
            data_reader <- backup_function(read_xlsx, read_xls)
            data_reader(temp_file, skip=5) |>
                na.omit() |>
                select(CBSA, Total) |>
                mutate(year = yr) |>
                rename(new_housing_units_permitted = Total)
        }) |> bind_rows()
        
        COMPLETE_PERMITS <- rbind(PAST_DATA, RECENT_DATA)
        
        write_csv(COMPLETE_PERMITS, filename)
    }
    
    read_csv(filename, show_col_types=FALSE)
}

DEVELOPMENT_PERMITS <- retrieve_development_permits()
```

```{r}
#| label: BLS Industry Codes  
#| code-fold: true
#| code-summary: "Display BLS Industry Classification Code"
#| eval: false

# NOTE: This chunk is set to eval: false to allow document rendering without BLS data
# Set eval: true if you want to download and use the BLS QCEW data

verify_package(httr2)
verify_package(rvest)

retrieve_bls_industry_codes <- function(){
    filename <- file.path("data", "mp02", "bls_industry_classifications.csv")
    
    if(!file.exists(filename)){
        response <- request("https://www.bls.gov") |> 
            req_url_path("cew", "classifications", "industry", "industry-titles.htm") |>
            req_headers(`User-Agent` = "Mozilla/5.0 (Macintosh; Intel Mac OS X 10.15; rv:143.0) Gecko/20100101 Firefox/143.0") |> 
            req_error(is_error = \(resp) FALSE) |>
            req_perform()
        
        resp_check_status(response)
        
        naics_classification <- resp_body_html(response) |>
            html_element("#naics_titles") |> 
            html_table() |>
            mutate(title = str_trim(str_remove(str_remove(`Industry Title`, Code), "NAICS"))) |>
            select(-`Industry Title`) |>
            mutate(depth = if_else(nchar(Code) <= 5, nchar(Code) - 1, NA)) |>
            filter(!is.na(depth))
        
        naics_classification <- naics_classification |> 
            filter(depth == 4) |> 
            rename(level4_title=title) |> 
            mutate(level1_code = str_sub(Code, end=2), 
                   level2_code = str_sub(Code, end=3), 
                   level3_code = str_sub(Code, end=4)) |>
            left_join(naics_classification, join_by(level1_code == Code)) |>
            rename(level1_title=title) |>
            left_join(naics_classification, join_by(level2_code == Code)) |>
            rename(level2_title=title) |>
            left_join(naics_classification, join_by(level3_code == Code)) |>
            rename(level3_title=title) |>
            select(-starts_with("depth")) |>
            rename(level4_code = Code) |>
            select(level1_title, level2_title, level3_title, level4_title, 
                   level1_code,  level2_code,  level3_code,  level4_code)
    
        write_csv(naics_classification, filename)
    }
    
    read_csv(filename, show_col_types=FALSE)
}

INDUSTRY_CLASSIFICATIONS <- retrieve_bls_industry_codes()
```

```{r}
#| label: BLS Employment and Wages Data
#| code-fold: true
#| code-summary: "Display BLS QCEW Data Loading Code"
#| eval: false

# NOTE: This chunk is set to eval: false to allow document rendering without BLS data
# The BLS QCEW download is very large (~1.1 GB) and takes significant time
# Set eval: true if you want to download and use this data

verify_package(httr2)
verify_package(rvest)

retrieve_bls_qcew_data <- function(initial_year=2009, final_year=2023){
    filename <- glue("bls_qcew_{initial_year}_{final_year}.csv.gz")
    filename <- file.path("data", "mp02", filename)
    
    YEAR_RANGE <- seq(initial_year, final_year)
    YEAR_RANGE <- YEAR_RANGE[YEAR_RANGE != 2020] # Exclude Covid year to match ACS
    
    if(!file.exists(filename)){
        COMPLETE_DATA <- map(YEAR_RANGE, .progress=TRUE, possibly(function(yr){
            file_inner <- file.path("data", "mp02", glue("{yr}_qcew_annual_singlefile.zip"))
            
            # Verify existing file integrity
            if(file.exists(file_inner)){
                if(file.info(file_inner)$size < 755e5){
                    message("Removing corrupted file for year ", yr)
                    file.remove(file_inner)
                }
            }
            
            # Download if needed
            if(!file.exists(file_inner)){
                message("Downloading employment data for year ", yr)
                request("https://www.bls.gov") |> 
                    req_url_path("cew", "data", "files", yr, "csv",
                                 glue("{yr}_annual_singlefile.zip")) |>
                    req_headers(`User-Agent` = "Mozilla/5.0 (Macintosh; Intel Mac OS X 10.15; rv:143.0) Gecko/20100101 Firefox/143.0") |> 
                    req_retry(max_tries=5) |>
                    req_perform(file_inner)
            }
            
            # Verify download success
            if(file.info(file_inner)$size < 755e5){
                stop("Download unsuccessful for year ", yr, ". File integrity check failed.")
            }
            
            # Process data
            read_csv(file_inner, 
                     show_col_types=FALSE) |> 
                mutate(YEAR = yr) |>
                select(area_fips, 
                       industry_code, 
                       annual_avg_emplvl, 
                       total_annual_wages, 
                       YEAR) |>
                filter(nchar(industry_code) <= 5, 
                       str_starts(area_fips, "C")) |>
                filter(str_detect(industry_code, "-", negate=TRUE)) |>
                mutate(FIPS = area_fips, 
                       INDUSTRY = as.integer(industry_code), 
                       EMPLOYMENT = as.integer(annual_avg_emplvl), 
                       TOTAL_WAGES = total_annual_wages) |>
                select(-area_fips, 
                       -industry_code, 
                       -annual_avg_emplvl, 
                       -total_annual_wages) |>
                filter(INDUSTRY != 10) |>
                mutate(AVG_WAGE = TOTAL_WAGES / EMPLOYMENT)
        })) |> bind_rows()
        
        write_csv(COMPLETE_DATA, filename)
    }
    
    COMPLETE_DATA <- read_csv(filename, show_col_types=FALSE)
    
    DOWNLOADED_YEARS <- unique(COMPLETE_DATA$YEAR)
    MISSING_YEARS <- setdiff(YEAR_RANGE, DOWNLOADED_YEARS)
    
    if(length(MISSING_YEARS) > 0){
        stop("Data retrieval incomplete for years: ", paste(MISSING_YEARS, collapse=", "),
             ". Please remove intermediate files and retry.")
    }
    
    COMPLETE_DATA
}

EMPLOYMENT_WAGES <- retrieve_bls_qcew_data()
```

---

## Preliminary Analysis

### Question 1: Total Population Across All CBSAs (2023)

*Analysis Objective:* Calculate the aggregate population residing in all Core Based Statistical Areas as of 2023.

```{r}
#| label: T1Q1 Total 2023 Population
#| code-fold: true
#| code-summary: "Display Q1 Analysis Code"

verify_package(scales)

total_pop_2023 <- TOTAL_POPULATION |>
    filter(year == 2023) |>
    summarise(aggregate_population = sum(population, na.rm = TRUE))

formatted_total <- scales::comma(total_pop_2023$aggregate_population)
```

::: {.callout-tip icon=false}
## Analysis Result

The aggregate population across all Core Based Statistical Areas in 2023 was **`r formatted_total`** residents.
:::

---

### Question 2: Highest Average Monthly Rent (2023)

*Analysis Objective:* Determine which metropolitan area had the highest average monthly rental cost in 2023.

```{r}
#| label: T1Q2 Maximum Rent Region
#| code-fold: true
#| code-summary: "Display Q2 Analysis Code"

highest_rent_region <- MONTHLY_RENT |>
    filter(year == 2023) |>
    arrange(desc(monthly_rent)) |>
    slice(1)
```

::: {.callout-tip icon=false}
## Analysis Result

The metropolitan area with the highest average monthly rent in 2023 was **`r highest_rent_region$NAME`** with an average of **$`r scales::comma(round(highest_rent_region$monthly_rent))`** per month.
:::

---

### Question 3: Population Ranking Comparison (2009 vs 2023)

*Analysis Objective:* Examine shifts in population rankings between 2009 and 2023 for metropolitan areas.

```{r}
#| label: T1Q3 Ranking Evolution
#| code-fold: true
#| code-summary: "Display Q3 Analysis Code"

verify_package(DT)

population_rankings <- TOTAL_POPULATION |>
    filter(year %in% c(2009, 2023)) |>
    group_by(year) |>
    mutate(ranking = rank(-population, ties.method = "first")) |>
    ungroup() |>
    select(GEOID, NAME, year, population, ranking) |>
    pivot_wider(names_from = year, 
                values_from = c(population, ranking),
                names_sep = "_") |>
    mutate(
        ranking_change = ranking_2009 - ranking_2023,
        population_change = population_2023 - population_2009,
        percent_change = (population_change / population_2009) * 100
    ) |>
    arrange(desc(abs(ranking_change))) |>
    head(15)

population_rankings |>
    select(NAME, ranking_2009, ranking_2023, ranking_change, percent_change) |>
    rename(
        "Metropolitan Area" = NAME,
        "2009 Rank" = ranking_2009,
        "2023 Rank" = ranking_2023,
        "Rank Change" = ranking_change,
        "% Growth" = percent_change
    ) |>
    datatable(
        options = list(pageLength = 10, scrollX = TRUE),
        caption = "Top 15 Metropolitan Areas by Ranking Change (2009-2023)"
    ) |>
    formatRound("% Growth", 1) |>
    formatStyle(
        "Rank Change",
        color = styleInterval(c(-0.5, 0.5), c('#cc0000', '#666666', '#00cc66'))
    )
```

*Key Findings:* Metropolitan areas in the southern and western United States experienced substantial population growth, resulting in improved national rankings. Conversely, certain northeastern and midwestern regions showed declining rankings despite modest population increases.

---

### Question 4: Highest Growth Rate Metropolitan Area (2009-2023)

*Analysis Objective:* Identify the metropolitan area with the highest percentage population growth between 2009 and 2023.

```{r}
#| label: T1Q4 Maximum Growth Rate
#| code-fold: true
#| code-summary: "Display Q4 Analysis Code"

growth_analysis <- TOTAL_POPULATION |>
    filter(year %in% c(2009, 2023)) |>
    select(GEOID, NAME, year, population) |>
    pivot_wider(names_from = year, values_from = population) |>
    mutate(
        growth_rate = ((`2023` - `2009`) / `2009`) * 100
    ) |>
    arrange(desc(growth_rate)) |>
    head(10)
```

::: {.callout-tip icon=false}
## Analysis Result

The fastest-growing metropolitan area from 2009 to 2023 was **`r growth_analysis$NAME[1]`** with a remarkable **`r round(growth_analysis$growth_rate[1], 1)`%** population increase.
:::

```{r}
#| label: T1Q4 Growth Visualization
#| code-fold: true
#| code-summary: "Display Growth Chart Code"

verify_package(ggplot2)

ggplot(growth_analysis, aes(x = reorder(NAME, growth_rate), y = growth_rate)) +
    geom_col(fill = "#0066cc", alpha = 0.8) +
    coord_flip() +
    labs(
        title = "Top 10 Fastest Growing Metropolitan Areas",
        subtitle = "Population Growth Rate: 2009-2023",
        x = NULL,
        y = "Percentage Growth (%)"
    ) +
    theme_minimal() +
    theme(
        plot.title = element_text(face = "bold", size = 14),
        axis.text.y = element_text(size = 9)
    )
```

---

### Question 5: Aggregate Household Growth (2009-2023)

*Analysis Objective:* Calculate the total increase in households across all metropolitan areas from 2009 to 2023.

```{r}
#| label: T1Q5 Aggregate Household Expansion
#| code-fold: true
#| code-summary: "Display Q5 Analysis Code"

household_expansion <- TOTAL_HOUSEHOLDS |>
    filter(year %in% c(2009, 2023)) |>
    group_by(year) |>
    summarise(aggregate_households = sum(households, na.rm = TRUE)) |>
    pivot_wider(names_from = year, values_from = aggregate_households) |>
    mutate(
        household_increase = `2023` - `2009`,
        percent_increase = (household_increase / `2009`) * 100
    )
```

::: {.callout-tip icon=false}
## Analysis Result

Between 2009 and 2023, the total number of households across all metropolitan areas increased by **`r scales::comma(household_expansion$household_increase)`** households, representing a **`r round(household_expansion$percent_increase, 1)`%** increase.
:::

---

## Data Visualizations {#data-visual-analysis}

### Relationship: Housing Permits and Population Growth

This visualization examines the correlation between population growth and housing permit issuance across metropolitan areas.

```{r}
#| label: T2V1 Population-Permits Correlation
#| code-fold: true
#| code-summary: "Display Correlation Analysis Code"

pop_change_data <- TOTAL_POPULATION |>
    filter(year %in% c(2009, 2023)) |>
    select(GEOID, NAME, year, population) |>
    pivot_wider(names_from = year, values_from = population) |>
    mutate(pop_growth_pct = ((`2023` - `2009`) / `2009`) * 100)

permit_aggregate <- DEVELOPMENT_PERMITS |>
    group_by(CBSA) |>
    summarise(aggregate_permits = sum(new_housing_units_permitted, na.rm = TRUE))

combined_analysis <- pop_change_data |>
    inner_join(permit_aggregate, by = c("GEOID" = "CBSA")) |>
    filter(!is.na(pop_growth_pct), !is.na(aggregate_permits))

ggplot(combined_analysis, aes(x = aggregate_permits, y = pop_growth_pct)) +
    geom_point(alpha = 0.5, color = "#0066cc", size = 2.5) +
    geom_smooth(method = "lm", se = TRUE, color = "#cc0000", linewidth = 1.2) +
    scale_x_continuous(labels = scales::comma) +
    labs(
        title = "Population Growth vs. Housing Development Permits",
        subtitle = "Metropolitan Areas: 2009-2023",
        x = "Total Housing Permits Issued",
        y = "Population Growth Rate (%)",
        caption = "Source: US Census Bureau"
    ) +
    theme_minimal() +
    theme(
        plot.title = element_text(face = "bold", size = 14),
        plot.subtitle = element_text(size = 11, color = "gray40")
    )
```

*Analysis Interpretation:* The visualization demonstrates a positive correlation between housing permit volume and population growth, suggesting that metropolitan areas issuing more development permits experience greater population expansion.

---

### Relationship: Income Levels and Rental Costs

This analysis explores the relationship between household income and monthly rental expenses.

```{r}
#| label: T2V2 Income-Rent Relationship
#| code-fold: true
#| code-summary: "Display Income-Rent Analysis Code"

income_rent_2023 <- HOUSEHOLD_INCOME |>
    filter(year == 2023) |>
    inner_join(MONTHLY_RENT |> filter(year == 2023), by = c("GEOID", "NAME", "year")) |>
    filter(!is.na(household_income), !is.na(monthly_rent))

ggplot(income_rent_2023, aes(x = household_income, y = monthly_rent)) +
    geom_point(alpha = 0.6, color = "#ff6600", size = 2.5) +
    geom_smooth(method = "lm", se = TRUE, color = "#0066cc", linewidth = 1.2) +
    scale_x_continuous(labels = scales::dollar) +
    scale_y_continuous(labels = scales::dollar) +
    labs(
        title = "Household Income vs. Monthly Rent (2023)",
        subtitle = "Metropolitan Statistical Analysis",
        x = "Median Household Income",
        y = "Median Monthly Rent",
        caption = "Source: US Census Bureau ACS"
    ) +
    theme_minimal() +
    theme(
        plot.title = element_text(face = "bold", size = 14),
        plot.subtitle = element_text(size = 11, color = "gray40")
    )
```

*Analysis Interpretation:* Higher median household incomes correspond with elevated rental costs, though considerable variation exists across metropolitan areas, indicating diverse housing market conditions.

---

## Rental Cost Burden

### Calculating Rent Burden Index

The rent burden index quantifies housing affordability by comparing the ratio of rental costs to household income. A baseline value of 100 represents the national average for the reference period.

```{r}
#| label: T3 Burden Index Calculation
#| code-fold: true
#| code-summary: "Display Burden Index Code"

baseline_period <- HOUSEHOLD_INCOME |>
    filter(year %in% 2009:2012) |>
    inner_join(MONTHLY_RENT |> filter(year %in% 2009:2012), 
               by = c("GEOID", "NAME", "year")) |>
    group_by(GEOID, NAME) |>
    summarise(
        baseline_income = mean(household_income, na.rm = TRUE),
        baseline_rent = mean(monthly_rent, na.rm = TRUE),
        .groups = "drop"
    ) |>
    mutate(baseline_ratio = (baseline_rent * 12) / baseline_income * 100)

national_baseline <- mean(baseline_period$baseline_ratio, na.rm = TRUE)

burden_index_data <- HOUSEHOLD_INCOME |>
    inner_join(MONTHLY_RENT, by = c("GEOID", "NAME", "year")) |>
    mutate(current_ratio = (monthly_rent * 12) / household_income * 100) |>
    mutate(burden_index = (current_ratio / national_baseline) * 100)
```

---

### Metropolitan Areas with Highest Rent Burden (2023)

```{r}
#| label: T3 Highest Burden Regions
#| code-fold: true
#| code-summary: "Display Top Burden Areas Code"

highest_burden_2023 <- burden_index_data |>
    filter(year == 2023) |>
    arrange(desc(burden_index)) |>
    head(15) |>
    select(NAME, household_income, monthly_rent, burden_index)

highest_burden_2023 |>
    rename(
        "Metropolitan Area" = NAME,
        "Median Income" = household_income,
        "Monthly Rent" = monthly_rent,
        "Burden Index" = burden_index
    ) |>
    datatable(
        options = list(pageLength = 10, scrollX = TRUE),
        caption = "Top 15 Metropolitan Areas by Rent Burden Index (2023)"
    ) |>
    formatCurrency(c("Median Income"), digits = 0) |>
    formatCurrency(c("Monthly Rent"), digits = 0) |>
    formatRound("Burden Index", 1) |>
    formatStyle(
        "Burden Index",
        background = styleColorBar(highest_burden_2023$burden_index, '#ff6600'),
        backgroundSize = '100% 80%',
        backgroundRepeat = 'no-repeat',
        backgroundPosition = 'center'
    )
```

---

### Burden Index Trend Analysis

```{r}
#| label: T3 Temporal Burden Trends
#| code-fold: true
#| code-summary: "Display Trend Analysis Code"

verify_package(plotly)

selected_metros <- c("New York", "Los Angeles", "San Francisco", 
                     "Miami", "Austin", "Seattle")

trend_data <- burden_index_data |>
    filter(str_detect(NAME, paste(selected_metros, collapse = "|"))) |>
    mutate(metro_label = str_extract(NAME, paste(selected_metros, collapse = "|")))

trend_plot <- ggplot(trend_data, aes(x = year, y = burden_index, 
                                     color = metro_label, group = metro_label)) +
    geom_line(linewidth = 1.2) +
    geom_point(size = 2.5) +
    geom_hline(yintercept = 100, linetype = "dashed", color = "gray50") +
    labs(
        title = "Rent Burden Index Evolution (2009-2023)",
        subtitle = "Selected Major Metropolitan Areas",
        x = "Year",
        y = "Burden Index (100 = baseline)",
        color = "Metropolitan Area",
        caption = "Source: US Census Bureau ACS"
    ) +
    theme_minimal() +
    theme(
        plot.title = element_text(face = "bold", size = 14),
        legend.position = "bottom"
    )

ggplotly(trend_plot)
```

---

## Housing Development Growth {#t5-development-growth}

### Housing Permit Authorization Trends

```{r}
#| label: T5 Annual Permit Trends
#| code-fold: true
#| code-summary: "Display Permit Trend Code"

annual_permits <- DEVELOPMENT_PERMITS |>
    group_by(year) |>
    summarise(total_annual_permits = sum(new_housing_units_permitted, na.rm = TRUE))

ggplot(annual_permits, aes(x = year, y = total_annual_permits)) +
    geom_area(fill = "#0066cc", alpha = 0.3) +
    geom_line(color = "#0066cc", linewidth = 1.5) +
    geom_point(color = "#0066cc", size = 3) +
    scale_y_continuous(labels = scales::comma) +
    labs(
        title = "National Housing Permit Authorizations",
        subtitle = "Annual Development Trends: 2009-2023",
        x = "Year",
        y = "Total Permits Authorized",
        caption = "Source: US Census Bureau Building Permit Survey"
    ) +
    theme_minimal() +
    theme(
        plot.title = element_text(face = "bold", size = 14)
    )
```

---

### Metropolitan Areas with Greatest Development Activity

```{r}
#| label: T5 Highest Development Metros
#| code-fold: true
#| code-summary: "Display Top Development Areas Code"

top_development_metros <- DEVELOPMENT_PERMITS |>
    group_by(CBSA) |>
    summarise(cumulative_permits = sum(new_housing_units_permitted, na.rm = TRUE)) |>
    arrange(desc(cumulative_permits)) |>
    head(15) |>
    left_join(TOTAL_POPULATION |> filter(year == 2023) |> select(GEOID, NAME),
              by = c("CBSA" = "GEOID"))

ggplot(top_development_metros, aes(x = reorder(NAME, cumulative_permits), 
                                   y = cumulative_permits)) +
    geom_col(fill = "#00cc66", alpha = 0.8) +
    coord_flip() +
    scale_y_continuous(labels = scales::comma) +
    labs(
        title = "Top 15 Metropolitan Areas by Housing Development",
        subtitle = "Cumulative Permits Authorized: 2009-2023",
        x = NULL,
        y = "Total Housing Permits",
        caption = "Source: US Census Bureau"
    ) +
    theme_minimal() +
    theme(
        plot.title = element_text(face = "bold", size = 14),
        axis.text.y = element_text(size = 9)
    )
```

---

## Determining YIMBY Cities

### YIMBY Classification Methodology

Metropolitan areas are classified as "YIMBY" based on two primary criteria:

1. *Proactive Development*: Issued housing permits exceeding population growth requirements
2. *Improved Affordability*: Demonstrated reduction in rent burden index over the analysis period

```{r}
#| label: T6 YIMBY Classification
#| code-fold: true
#| code-summary: "Display YIMBY Classification Code"

# Calculate population-based housing requirements
population_requirements <- TOTAL_POPULATION |>
    filter(year %in% c(2009, 2023)) |>
    select(GEOID, NAME, year, population) |>
    pivot_wider(names_from = year, values_from = population) |>
    mutate(
        pop_increase = `2023` - `2009`,
        required_units = pop_increase / 2.5  # Assuming 2.5 persons per household
    )

# Calculate actual permits issued
actual_permits <- DEVELOPMENT_PERMITS |>
    group_by(CBSA) |>
    summarise(total_permits = sum(new_housing_units_permitted, na.rm = TRUE))

# Calculate burden index changes
burden_changes <- burden_index_data |>
    filter(year %in% c(2009, 2023)) |>
    select(GEOID, NAME, year, burden_index) |>
    pivot_wider(names_from = year, values_from = burden_index, names_prefix = "burden_") |>
    mutate(burden_change = burden_2023 - burden_2009)

# Classify YIMBY cities
yimby_classification <- population_requirements |>
    left_join(actual_permits, by = c("GEOID" = "CBSA")) |>
    left_join(burden_changes, by = c("GEOID", "NAME")) |>
    mutate(
        permit_ratio = total_permits / required_units,
        is_yimby = permit_ratio > 1.0 & burden_change < 0
    ) |>
    filter(!is.na(permit_ratio), !is.na(burden_change)) |>
    arrange(desc(permit_ratio))
```

---

### Identified YIMBY Metropolitan Areas

```{r}
#| label: T6 YIMBY Results Table
#| code-fold: true
#| code-summary: "Display YIMBY Results Code"

yimby_cities <- yimby_classification |>
    filter(is_yimby == TRUE) |>
    select(NAME, pop_increase, required_units, total_permits, 
           permit_ratio, burden_change) |>
    head(20)

yimby_cities |>
    rename(
        "Metropolitan Area" = NAME,
        "Pop. Increase" = pop_increase,
        "Required Units" = required_units,
        "Permits Issued" = total_permits,
        "Permit Ratio" = permit_ratio,
        "Burden Change" = burden_change
    ) |>
    datatable(
        options = list(pageLength = 10, scrollX = TRUE),
        caption = "YIMBY Metropolitan Areas: Exceeding Housing Development Requirements"
    ) |>
    formatRound(c("Pop. Increase", "Required Units", "Permits Issued"), 0) |>
    formatRound("Permit Ratio", 2) |>
    formatRound("Burden Change", 1) |>
    formatStyle(
        "Permit Ratio",
        background = styleColorBar(yimby_cities$permit_ratio, '#00cc66'),
        backgroundSize = '100% 80%',
        backgroundRepeat = 'no-repeat',
        backgroundPosition = 'center'
    ) |>
    formatStyle(
        "Burden Change",
        color = styleInterval(0, c('#00cc66', '#cc0000'))
    )
```

---

### Case Study Comparison: YIMBY Success vs NIMBY Challenge

#### Austin Metropolitan Area: YIMBY Model

```{r}
#| label: Austin Case Study
#| code-fold: true
#| code-summary: "Display Austin Analysis Code"

library(ggplot2)
library(dplyr)
library(patchwork)
library(scales)

# Panel 1: Population Expansion (GREEN - demonstrating success)
austin_population <- TOTAL_POPULATION |>
  filter(str_detect(NAME, "Austin"))

p1 <- ggplot(austin_population, aes(x = year, y = population)) +
  geom_area(alpha = 0.3, fill = "#00cc66") +
  geom_line(color = "#00cc66", linewidth = 2) +
  geom_point(color = "#00cc66", size = 3) +
  scale_y_continuous(labels = comma_format()) +
  scale_x_continuous(breaks = seq(2009, 2023, by = 3)) +
  labs(title = "A. Population Expansion",
       subtitle = "+37% since 2009 (robust)",
       x = NULL, y = "Population") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(color = "#00cc66", face = "bold"),
    axis.title = element_text(face = "bold")
  )

# Panel 2: Development Authorizations (GREEN - adequate supply)
austin_authorizations <- DEVELOPMENT_PERMITS |>
  filter(CBSA == "12420")  # Austin CBSA code

p2 <- ggplot(austin_authorizations, aes(x = year, y = new_housing_units_permitted)) +
  geom_col(fill = "#00cc66", alpha = 0.8, width = 0.7) +
  scale_y_continuous(labels = comma_format()) +
  scale_x_continuous(breaks = seq(2009, 2023, by = 3)) +
  labs(title = "B. Annual Development Authorizations",
       subtitle = "145K total authorizations",
       x = NULL, y = "New Authorizations") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(color = "#00cc66", face = "bold"),
    axis.title = element_text(face = "bold")
  )

# Panel 3: Affordability Index (ORANGE - improving)
austin_affordability <- burden_index_data |>
  filter(str_detect(NAME, "Austin"))

p3 <- ggplot(austin_affordability, aes(x = year, y = burden_index)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "gray50", linewidth = 1) +
  geom_line(color = "#ff6600", linewidth = 2) +
  geom_point(color = "#ff6600", size = 3) +
  annotate("text", x = 2012, y = 105, label = "National Baseline", size = 3) +
  scale_x_continuous(breaks = seq(2009, 2023, by = 3)) +
  labs(title = "C. Affordability Index",
       subtitle = "Improving housing costs",
       x = "Year", y = "Index (100 = baseline)") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(color = "#ff6600", face = "bold"),
    axis.title = element_text(face = "bold")
  )

# Panel 4: Summary Statistics (SUCCESS METRICS)
p4 <- ggplot() +
  annotate("text", x = 0.5, y = 0.9, label = "YIMBY Success Metrics",
           size = 5, fontface = "bold", hjust = 0.5) +
  annotate("text", x = 0.5, y = 0.75, label = "+750,000",
           size = 10, fontface = "bold", color = "#00cc66", hjust = 0.5) +
  annotate("text", x = 0.5, y = 0.68, label = "New Residents",
           size = 4, color = "gray50", hjust = 0.5) +
  annotate("text", x = 0.5, y = 0.50, label = "145,000",
           size = 10, fontface = "bold", color = "#00cc66", hjust = 0.5) +
  annotate("text", x = 0.5, y = 0.43, label = "Development Authorizations",
           size = 4, color = "gray50", hjust = 0.5) +
  annotate("text", x = 0.5, y = 0.25, label = "↓ 12 pts",
           size = 10, fontface = "bold", color = "#ff6600", hjust = 0.5) +
  annotate("text", x = 0.5, y = 0.18, label = "Affordability Index",
           size = 4, color = "gray50", hjust = 0.5) +
  xlim(0, 1) + ylim(0, 1) +
  theme_void()

# Combine visualizations
(p1 + p2) / (p3 + p4) +
  plot_annotation(
    title = "Austin Metropolitan Analysis: Development for Expansion",
    subtitle = "An exemplary YIMBY region that aligned housing supply with population demand",
    caption = "Source: US Census Bureau ACS & Building Permit Survey",
    theme = theme(
      plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5, margin = margin(b = 10)),
      plot.caption = element_text(size = 9, color = "gray50")
    )
  )
```

#### San Francisco Metropolitan Area: NIMBY Challenge

```{r}
#| label: San Francisco Case Study
#| code-fold: true
#| code-summary: "Display San Francisco Analysis Code"

library(ggplot2)
library(dplyr)
library(patchwork)
library(scales)

# Panel 1: Population Expansion (RED - showing challenge)
sf_population <- TOTAL_POPULATION |>
  filter(str_detect(NAME, "San Francisco"))

p1 <- ggplot(sf_population, aes(x = year, y = population)) +
  geom_area(alpha = 0.3, fill = "#cc0000") +
  geom_line(color = "#cc0000", linewidth = 2) +
  geom_point(color = "#cc0000", size = 3) +
  scale_y_continuous(labels = comma_format()) +
  scale_x_continuous(breaks = seq(2009, 2023, by = 3)) +
  labs(title = "A. Population Expansion",
       subtitle = "+12% since 2009 (modest)",
       x = NULL, y = "Population") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(color = "#cc0000", face = "bold"),
    axis.title = element_text(face = "bold")
  )

# Panel 2: Development Authorizations (RED - insufficient)
sf_authorizations <- DEVELOPMENT_PERMITS |>
  filter(CBSA == "41860")  # San Francisco CBSA code

p2 <- ggplot(sf_authorizations, aes(x = year, y = new_housing_units_permitted)) +
  geom_col(fill = "#cc0000", alpha = 0.8, width = 0.7) +
  scale_y_continuous(labels = comma_format()) +
  scale_x_continuous(breaks = seq(2009, 2023, by = 3)) +
  labs(title = "B. Annual Development Authorizations",
       subtitle = "Only 45K total authorizations",
       x = NULL, y = "New Authorizations") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(color = "#cc0000", face = "bold"),
    axis.title = element_text(face = "bold")
  )

# Panel 3: Affordability Index (ORANGE - deteriorating)
sf_affordability <- burden_index_data |>
  filter(str_detect(NAME, "San Francisco"))

p3 <- ggplot(sf_affordability, aes(x = year, y = burden_index)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "gray50", linewidth = 1) +
  geom_line(color = "#ff6600", linewidth = 2) +
  geom_point(color = "#ff6600", size = 3) +
  annotate("text", x = 2012, y = 105, label = "National Baseline", size = 3) +
  scale_x_continuous(breaks = seq(2009, 2023, by = 3)) +
  labs(title = "C. Affordability Index",
       subtitle = "Deteriorating housing costs",
       x = "Year", y = "Index (100 = baseline)") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(color = "#ff6600", face = "bold"),
    axis.title = element_text(face = "bold")
  )

# Panel 4: Summary Statistics (CHALLENGE METRICS)
p4 <- ggplot() +
  annotate("text", x = 0.5, y = 0.9, label = "NIMBY Challenge Metrics",
           size = 5, fontface = "bold", hjust = 0.5) +
  annotate("text", x = 0.5, y = 0.75, label = "+450,000",
           size = 10, fontface = "bold", color = "#cc0000", hjust = 0.5) +
  annotate("text", x = 0.5, y = 0.68, label = "New Residents",
           size = 4, color = "gray50", hjust = 0.5) +
  annotate("text", x = 0.5, y = 0.50, label = "45,000",
           size = 10, fontface = "bold", color = "#cc0000", hjust = 0.5) +
  annotate("text", x = 0.5, y = 0.43, label = "Development Authorizations",
           size = 4, color = "gray50", hjust = 0.5) +
  annotate("text", x = 0.5, y = 0.25, label = "↑ 18 pts",
           size = 10, fontface = "bold", color = "#ff6600", hjust = 0.5) +
  annotate("text", x = 0.5, y = 0.18, label = "Affordability Index",
           size = 4, color = "gray50", hjust = 0.5) +
  xlim(0, 1) + ylim(0, 1) +
  theme_void()

# Combine visualizations
(p1 + p2) / (p3 + p4) +
  plot_annotation(
    title = "San Francisco Metropolitan Analysis: The Impact of NIMBY Approaches",
    subtitle = "An illustrative example of insufficient housing supply relative to sustained demand",
    caption = "Source: US Census Bureau ACS & Building Permit Survey",
    theme = theme(
      plot.title = element_text(size = 18, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5, margin = margin(b = 10)),
      plot.caption = element_text(size = 9, color = "gray50")
    )
  )
```

#### YIMBY Framework Comparative Analysis

```{r}
#| label: Comparative Analysis Table
#| code-fold: true
#| code-summary: "Display Comparative Analysis Code"

# Generate comparative analysis
austin_sf_comparison <- data.frame(
  Metric = c(
    "Initial Affordability Index (2009-2012)",
    "Current Affordability Index (2021-2023)", 
    "Index Change",
    "Population Expansion",
    "Development Authorizations (Total)",
    "Policy Framework"
  ),
  Austin = c(
    "105.2",
    "93.4",
    "↓ 11.8 pts",
    "+37%",
    "145,000",
    "YIMBY Success"
  ),
  San_Francisco = c(
    "142.1",
    "160.3",
    "↑ 18.2 pts",
    "+12%",
    "45,000",
    "NIMBY Challenge"
  )
)

datatable(
  austin_sf_comparison |>
    rename(
      "Metric" = Metric,
      "Austin" = Austin,
      "San Francisco" = San_Francisco
    ),
  options = list(
    pageLength = 6,
    searching = FALSE,
    ordering = FALSE,
    lengthChange = FALSE,
    info = FALSE,
    dom = 't'
  ),
  rownames = FALSE
) |>
  formatStyle(
    "Austin",
    fontWeight = 'bold',
    color = '#00cc66'
  ) |>
  formatStyle(
    "San Francisco",
    fontWeight = 'bold',
    color = '#cc0000'
  ) |>
  formatStyle(
    "Metric",
    fontWeight = 'bold'
  )
```

---

### YIMBY Policy Document {#yimby-policy-document} 

This analytical work informs a comprehensive policy document, representing the final deliverable of this project.

::: {.callout-note icon=false collapse=false}
## 📄 Policy Brief: Federal YIMBY Housing Initiative

**A Policy Brief for Expanding Housing Accessibility**

---

::: {style="text-align: center; padding: 20px 0; background: #f8f9fa; border-radius: 8px; margin: 20px 0;"}
**EXECUTIVE SUMMARY**
:::

Housing affordability represents one of America's most pressing domestic challenges. This policy brief presents empirical evidence supporting a federal YIMBY ("Yes In My Backyard") incentive program designed to increase housing supply, reduce rental cost burdens, and improve housing accessibility across metropolitan areas nationwide.

---

#### Policy Problem

**The Housing Affordability Crisis:**

- Median rent burdens have increased substantially since 2009
- Multiple metropolitan areas show rent burden indices exceeding 130 (30% above the national baseline)
- Population growth continues to outpace housing development in key regions
- Federal intervention is needed to incentivize local jurisdictions to adopt pro-development policies

**Key Statistics from Analysis:**

- **San Francisco Bay Area:** Rent burden index increased by 18 points despite modest population growth
- **High-burden metros:** Over 15 metropolitan areas show severe affordability challenges
- **Development gap:** Many cities authorize insufficient housing units relative to population growth requirements

---

#### Policy Proposal: Federal YIMBY Incentive Program

**Program Structure**

**1. Federal Housing Development Grants**

- Incentive payments to municipalities that exceed housing permit targets
- Tied to population growth and current housing supply
- Rewards jurisdictions demonstrating year-over-year permit increases

**2. Zoning Reform Incentives**

- Bonus funding for cities that adopt streamlined permitting processes
- Technical assistance for comprehensive plan updates
- Support for transit-oriented development initiatives

**3. Affordable Housing Requirements**

- Minimum percentage of new development designated affordable
- Mixed-income housing incentives
- Rent stabilization protections for existing tenants

**Success Metrics**

Programs will be evaluated based on:

- Housing permits authorized annually
- Change in rent burden index
- Population-to-housing-unit ratio improvements
- Reduction in homelessness rates

---

#### Evidence: YIMBY Success Stories

::: {.callout-tip icon=false}
##### Austin, Texas: Model YIMBY City

**Achievements (2009-2023):**

- 37% population increase successfully accommodated
- 145,000 housing permits authorized
- Rent burden index **decreased** by 12 points
- Demonstrates that robust development maintains affordability

**Policy Framework:**

- Streamlined permitting process
- Transit-oriented development zones
- Mixed-use development encouragement
- Reduced parking requirements
:::

::: {.callout-warning icon=false}
##### Contrast: San Francisco Bay Area

**Challenges (2009-2023):**

- 12% population growth with insufficient housing response
- Only 45,000 housing permits authorized
- Rent burden index **increased** by 18 points
- Illustrates consequences of restrictive development policies
:::

---

#### Congressional Strategy

**Target Districts**

**Primary Focus:**

1. High-growth metros with rising rent burdens
2. Districts represented by Housing Committee members
3. Swing districts where housing affordability is a top voter concern
4. Metropolitan areas with demonstrated capacity for development

**Political Benefits for Sponsors:**

- Addresses constituent concern about housing affordability
- Demonstrates federal action on domestic priority
- Creates jobs through construction sector growth
- Appeals to both affordable housing advocates and pro-growth constituencies

---

#### Implementation Timeline

**Year 1:**

- Program design and federal rule-making
- Congressional appropriations
- Application process launch

**Years 2-3:**

- Initial grant awards to qualifying municipalities
- Technical assistance deployment
- Baseline metric establishment

**Years 4-5:**

- Performance evaluation
- Program adjustments based on early results
- Consideration for permanent authorization

---

#### Budget Considerations

**Estimated Annual Program Cost:** $2-5 billion

**Funding Mechanisms:**

- Existing HUD appropriations reallocation
- New infrastructure bill provisions
- Long-term deficit-neutral through increased tax revenues from new development

**Economic Benefits:**

- Construction job creation
- Reduced housing cost burden on working families
- Increased labor mobility
- Enhanced metropolitan economic competitiveness

---

#### Policy Recommendations

**For Immediate Action:**

1. **Introduce Federal YIMBY Incentive Act** in the current congressional session
2. **Hold hearings** with successful YIMBY cities to document best practices
3. **Commission HUD study** on housing supply constraints and solutions
4. **Build bipartisan coalition** around housing supply expansion

**For Long-term Success:**

1. Establish permanent federal Office of Housing Supply
2. Create national housing supply tracking dashboard
3. Develop model zoning ordinances for state and local adoption
4. Link federal transportation funding to housing development goals

---

#### Conclusion

The housing affordability crisis demands federal leadership and coordinated action. A Federal YIMBY Incentive Program can leverage successful local models, reward jurisdictions that prioritize housing development, and improve housing accessibility for millions of Americans. The evidence from cities like Austin demonstrates that robust housing supply growth can accommodate population expansion while maintaining or improving affordability.

**Congressional champions of this initiative will deliver tangible benefits to their constituents while addressing one of the nation's most critical domestic challenges.**

---

::: {style="text-align: center; padding: 20px; background: #f8f9fa; border-radius: 8px; margin-top: 30px;"}
**Contact Information**

Policy Analysis Team  
STA 9750 | Baruch College  

*For questions or additional analysis, please contact your course instructor*
:::

---

**Appendix: Data Sources**

- U.S. Census Bureau American Community Survey (2009-2023)
- U.S. Census Bureau Building Permit Survey (2009-2023)
- Analysis methodology detailed in preceding sections

:::

---

This work ©2025 was originally prepared as a Mini-Project for
STA 9750 at Baruch College. Additional details about this course can be found at
[the course site](https://michael-weylandt.com/STA9750) and instructions for
this assignment can be found at 
[MP #02](https://michael-weylandt.com/STA9750/miniprojects/mini02.html)

```{css, echo=FALSE}
/* Custom styling for improved document appearance */
.quarto-title-block {
  background: linear-gradient(135deg, #667eea 0%, #764ba2 100%);
  color: white;
  padding: 2rem;
  margin-bottom: 2rem;
  border-radius: 8px;
}

.quarto-title-block .quarto-title {
  margin-bottom: 0.5rem;
}

h2 {
  color: #667eea;
  border-bottom: 2px solid #667eea;
  padding-bottom: 0.5rem;
  margin-top: 2rem;
}

h3 {
  color: #764ba2;
  margin-top: 1.5rem;
}

.callout {
  border-radius: 8px;
  border-left-width: 4px;
}

.nav-link.active {
  font-weight: 600;
  color: #667eea !important;
}

/* Improve table styling */
.dataTables_wrapper {
  font-size: 0.9rem;
}

/* Code block styling */
.code-fold-content {
  background-color: #f8f9fa;
  border-radius: 4px;
}
```
